### 2021-05-23 DeepLearningAndBayesianSoftware.Rmd
getting close to release

---
title: "Deep Learning and Bayesian Modeling"
subtitle: "Comparing research use via citation counting"
author: "Breck Baldwin"
email: "breckbaldwin@gmail.com"
date: "5/23/2021"
output:
 bookdown::html_document2:
   includes:
     in_header: _html/ga.html 
bibliography: citations.bib
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	comment = NA
)
```

# Introduction

Stan, a Bayesian modeling language, was released in 2012 to considerable fanfare. A shiny new inference algorithm, HMC with NUTS (@hoffman2014no), promised and delivered fits that could not be fit before. And then deep learning happened. The relese of TensorFlow (@abadi2016tensorflow) opened the doors of deep learning to everyone and in 2016 began garnering significant traction. One result of this was people often assumed that deep learning's successes usurped Bayesian modeling's domain. This was not in our imagination, NSF reviews came back dismissing Stan funding because all the interesting work was assumed to be happening with deep learning. Mind you, deep learning and Bayesian modeling are conceptual cousins but very different from each other and are better thought of as complementary than as antagonistic. Yet Bayesians found themselves in deep learning's shadow somehow. It is 2021 and this document does some very simple analysis around use of Bayesian and deep learning packages as evidenced in the research literature to get a perspective on what actually happened and is happening. The comparison aims to approach the following goals with very simple research citation metrics:

1. How does Bayesian modeling software stack up against deep learning? 
2. Asses the impact of Bayesian modeling software using deep learning metrics as a yard stick--how big a fraction of a huge thing are we?
3. Contextualize the roles each approach has by looking at subject distributions. The technologies have very different use cases so one would expect variation.

Citation counting is a crude metric but it has the advantage of simplicity. In compiling these metrics I came away with very different opinion than I started with so I thought it worth sharing. My prior assumptions were that Bayesian modeling was very niche and scurrying around doing very useful and important science but niche none the less. This analysis led me to revise that opinion considerably.

For the purposes of this document we take Bayesian software to be Stan (@standev) and PyMC3 (@salvatier2016probabilistic) with ecosystem components included that are likely to be cited. That includes the simplified syntax interfaces to Stan, brms (@brms), RstanArm (@rstanarm), and interface packages to Stan, RStan (@rstan) and PyStan (@pystan)[^1]. The analysis applies only to packages that are in current development so the venerated, and very high impact, packages like BUGS (@lunn2013bugs) and JAGS (@plummer2003jags) are not considered although including them would double our counts. Also, the ecosystems are actually much larger but they are unlikely to be cited in the research literature and we have to stop some place. For deep learning we take TensorFlow, its interface Keras (@keras) and PyTorch (@pytorch) as roughly equivalent entities for the comparison. Theano (@theano) is no longer in development. It should be noted that both PyTorch and TensorFlow have implemented HMC with NUTS for Bayesian inference but those are recent developments that have not made much of an impact yet.

[^1]:There are also lighter weight interfaces as well interface to other languages that I ignore due to low expected research mentions, see [https://https://mc-stan.org/users/interfaces/]([https://https://mc-stan.org/users/interfaces/) for a listing.

The key resource behind this document is Elsevier's https://scopus.com research search engine that provides a tightly curated[^2] search index that includes sources outside of the academic behemoth's own journals. It also provides a solid API (Application Programmer Interface) and a classification of journals into subject areas. The actual form of Scopus queries is discussed below.

[^2]:https://scolar.google.com genrally yields double the counts for similar queries but no subject classification or API to code against. The increased counts are likely due to inclusion of non-peer reviewed sites like [https://arxiv.org](https://arxiv.org)

# High level prevalence of Deep Learning vs Bayesian Modeling

In the process of grant writing I create metrics to help justify projects, lately in partnership with PyMC and ArviZ through the scientific fiscal sponsor NumFOCUS of which Stan is a member as well. Since Bayesian modeling seems always in the shadow of deep learning I started tracking those software packages as well for comparison. Below we see the relative citations of the top deep learning packages TensorFlow, PyTorch and the support package Keras to the Bayesian packages PyMC3, Stan with support/derivative packages RStan, RStanArm, PyStan and brms.

```{r Access Scopus, include=FALSE}
rm(list = rm()) #clean out any state from previous runs
library(tidyverse)
library(jsonlite)
library(httr)
library(kableExtra)
library(stringr)

# remember to set working directory to the location of this file
setwd("~/git/ScientificSoftwareImpactMetrics")
credentials <- read_json('Scopus_credentials.json')
API_KEY <- credentials$API_KEY
INSTITUTION_TOKEN <- credentials$INSTITUTION_TOKEN
# Format of Scopus_credentials.json
# {
# "API_KEY":"XXXXXXXXXXXXXXXXXXXXXXXXXxx",
# "INSTITUTION_TOKEN":"XXXXXXXXXXXXXXXXXXXXXXXX"
# }

BASE_URL = 'https://api.elsevier.com/content/search/scopus'

USE_CACHE = TRUE #will have to setup a redis server, strongly recommend doing this
REPORT_PROGRESS = FALSE

if (USE_CACHE) {
  library(redux)
  redis <- redux::hiredis()
  get_results <- function(url) {
    cache <- redis$GET(url) # check redis first
    if (!is.null(cache)) {
      result <- unserialize(cache)
      if (result$status_code == 200) {
        if (REPORT_PROGRESS) {
          cat(paste("\nhitting redis for", url))
        }
        return(result)
      }
    }
    random_wait <- abs(rnorm(1, 1, 1))
    if (REPORT_PROGRESS) {
      cat(paste("\ncache miss, querying:", url, "\n"))
      cat(paste(
        "\nWaiting",
        random_wait,
        "seconds to be nice to webserver\n"
      ))
    }
    Sys.sleep(random_wait)
    result <- GET(
      url,
      add_headers('X-ELS-APIKey' = API_KEY,
                  'X-ELS-Insttoken' = INSTITUTION_TOKEN)
    )
    redis$SET(url, serialize(result, NULL))
    return(result)
  }
} else { #not using cache, hacky but works with R
  get_results <- function(url) {
    random_wait <- abs(rnorm(1, 1, 1))
    if (REPORT_PROGRESS) {
      cat(paste("\nno cache used, querying:", url, "\n"))
      cat(paste(
        "\nWaiting",
        random_wait,
        "seconds to be nice to webserver\n"
      ))
    }
    Sys.sleep(random_wait)
    result <- GET(
      url,
      add_headers('X-ELS-APIKey' = API_KEY,
                  'X-ELS-Insttoken' = INSTITUTION_TOKEN)
    )
    return(result)
  }
}

# setup queries for analysis

stan_q <- '(gelman+AND+hoffman+AND+stan)+OR+mc-stan.org'
pymc_q <- 'PyMC3+OR+(PyMC*+AND+fonnesbeck)'
rstan_q <- '(rstan+AND+NOT+mit)'
brms_q <-  '(brms+AND+burkner)'

#Display names followed by queries
pkg_query <- c('Stan', 'brms','PyStan', 'RStanArm', 'RStan','PyMC',  'TensorFlow', 'PyTorch', 'Keras',
               stan_q, brms_q, 'pystan', 'rstanarm', rstan_q, pymc_q, 'tensorflow', 'pytorch', 'Keras')
pkg_query_m <- matrix(data = pkg_query, ncol =2)
queryBayVsDeepM <- matrix(ncol = 2, nrow = 2)
queryBayVsDeepM[1,] <- c('Bayesian', paste(pkg_query_m[1:6,2], collapse = "+OR+"))
queryBayVsDeepM[2,] <- c('Deep Learning', paste(pkg_query_m[7:9,2], collapse = "+OR+"))

```

```{r DL vs Bayes Graph, echo=FALSE, message=FALSE, warning=FALSE}
library(ggrepel)
year_start <- 2012
year_end <- 2020
years <- year_start:year_end
df <- data.frame(years)

years <- year_start:year_end
scopus.df <- data.frame(years)
for (i in 1:nrow(queryBayVsDeepM)) {
  package_name = queryBayVsDeepM[i, 1]
  query = queryBayVsDeepM[i, 2]
  if (query == '') {
    query = package_name
  }
  for (suffix in c('','_no_comp_sci')) {
    year_counts = c()
        for (year in year_start:year_end) {
      url <- paste(BASE_URL, '?query=','REF(', query,')',
                   "+AND+PUBYEAR+=+", year,
                   sep = '')
      if (suffix == '_no_comp_sci') {
        url = paste(url, '+AND+NOT+SUBJAREA(COMP)', sep = '')
      }
      result <- get_results(url)
      json_txt <- rawToChar(as.raw(strtoi(result$content, 16L)))
      data <- jsonlite::fromJSON(json_txt)
      year_counts = c(
        year_counts,
        as.numeric(data$`search-results`$`opensearch:totalResults`))
    }
    scopus.df[paste(
      package_name,
      suffix,
      sep = '')] <- year_counts
  }
}

column_names <- colnames(scopus.df)
col_sums <- colSums(scopus.df)
df_long <- gather(scopus.df, topic, yr_count,
                  column_names[2]:column_names[length(column_names)])

df_long_label <- df_long %>%
  mutate(label = if_else(years == max(years),
                         paste(as.character(topic),col_sums[topic], sep = ':'), NA_character_))

plot2 <-
  ggplot(data = df_long_label, aes(
    x = years,
    y = yr_count,
    group = topic,
    color = topic
  )) +
  geom_line() +
  geom_point() +
  geom_label_repel(aes(label = label),
                   na.rm = TRUE) +
  scale_color_discrete(guide = FALSE) +
  labs(y = "year count for research articles matching topic",
  caption = "Fig 1: Relative counts with and without computer science journals") +
   theme(plot.caption=element_text(size=12, hjust=0, margin=margin(15,0,0,0)))

print(plot2)
```

Reading from top of Fig 1, assuming raw research citations counts matter, Bayesian packages are overwhelmed by deep learning packages with a count of nearly 35,000 citations to a Bayesian count of 4,500--nearly a 10x difference. One could excuse a bit of Bayesian despondency in the face of the wave depicted above--this is ratio mirrors my expectations. But graph also shows relative counts for research citations outside of computer science which drops the count of deep learning from 34,977 to 8,885--a 75% drop. That drop IS a surprise.  I didn't have a prior on deep learning's distribution across subject categories but I didn't expect it to be so skewed. The Bayesian ecosystem shows a total of 4,578 of which 14% are computer science--that drop does not surprise me.

An early take away is that Bayesian's shouldn't bother applying to computer science directorates since it's deep learning's home turf or Bayesians should double down on applying to computer science directorates since they must be getting sick of deep learning proposals.

# Subject Category Breakdowns

So what is going on here? The next question is what does the the distribution look like for non-computer science subject areas?  Below is a table of subject counts for packages starting in 2006 when PyMC was first released to 2021 with relative percentages listed:

```{r Table 1, echo=FALSE, message=FALSE, warning=FALSE}
library(scales)
library(kableExtra)

makeDisplayTable <- function(df) {
  subjCats = colnames(df)[-1:-2]
  df %>% group_by(package) %>% 
    summarize_each(funs(sum), all_of(subjCats)) -> outDf
  finalDf = t(outDf)
  subFrame = finalDf[-1,]
  cleanDf =  as.data.frame(apply(subFrame, 2, as.numeric))
  colnames(cleanDf) = finalDf[1,]
  rownames(cleanDf) = rownames(finalDf)[-1]
  cleanDf %>% mutate(totals = rowSums(.)) -> summedDf
  displayDf = summedDf
  displayDf[] = NA_character_
  for (i in 1:nrow(summedDf)) {
    displayDf[i,] = as.list(sprintf("%d/%s", as.numeric(summedDf[i,]), 
                                    label_percent(accuracy = 1)(as.numeric(summedDf[i,]/summedDf[i,]$totals))))
    displayDf[i,]$totals = summedDf[i,]$totals 
  }
  displayOrderedDf <- displayDf[order(row.names(displayDf)),] 
  return(displayOrderedDf)
}

getAllSubjectCategories <- function (queryM) {
  year_start = 2006
  year_end = 2021
  years <- year_start:year_end
  accumDf = data.frame(matrix(ncol = 2, nrow = nrow(queryM) * length(years)))
  colnames(accumDf) = c('package', 'years')
  counter = 1
  article_count = 0
  article_subj_count = 0
  for (i in 1:nrow(queryM)) {
    package_name = queryM[i, 1]
    query = queryM[i, 2]
    if (!is.na(str_match(query,"^\\(")) && !is.na(str_match(query,"\\)$"))) {
      query = str_match(query, "^\\((.*)\\)$")[1,2] #remove extra brackets, breaks scopus
    }
    accumDf[counter:((counter - 1) + length(years)),]$years <- years
    accumDf[counter:((counter - 1) + length(years)),]$package <- 
      rep(package_name, length(years))
    for (year in year_start:year_end) {
      url <-
        paste(
          BASE_URL,
          '?query=',
          'REF(',
          query,
          ')',
          "+AND+PUBYEAR+=+",
          year,
          '&facets=subjarea(count=101)',
          sep = ''
        )
      result <- get_results(url)
      if (result$status_code != 200) {
        print(sprintf("got non 200 status from query: %d", result$status_code))
        stop()
      }
      json_txt <- rawToChar(as.raw(strtoi(result$content, 16L)))
      data <- jsonlite::fromJSON(json_txt)
      article_count <-
        as.numeric(data$`search-results`$`opensearch:totalResults`) + article_count
      facet_count <- length(data$`search-results`$facet$category$name)
      j <- 1
      while (j < facet_count) {
        name <- data$`search-results`$facet$category$label[j]
        name <- str_replace(name, " \\(all\\)", "")
        hitCount <- as.numeric(data$`search-results`$facet$category$hitCount[j])
        article_subj_count <- hitCount + article_subj_count
        if (!name %in% colnames(accumDf)) {
          accumDf[name] <- rep(0, length(years) * nrow(queryM))
          #print(paste("name=",name,", count=",hitCount))
        }
        accumDf[counter,][[name]] <- hitCount
        j <- j+ 1
      }
      counter = counter + 1
    }
  }
  if (REPORT_PROGRESS) {
    print(sprintf("%d paperXsubject with %d total papers for %.3f avg subjects per paper",
                  article_subj_count,
                  article_count,
                  article_subj_count/article_count))
  }
  return(accumDf)
}

bayesVdeepLearnDf <- getAllSubjectCategories(queryBayVsDeepM)
displayBayesVdeepLearnDf <- makeDisplayTable(bayesVdeepLearnDf)

detailDf  <- getAllSubjectCategories(pkg_query_m)
displayDetailDf <- makeDisplayTable(detailDf)

allDataDf <- cbind(displayBayesVdeepLearnDf,displayDetailDf)
allDataDf <- allDataDf[,c(1,2,3,11,6,7,10,9,4,8,5,12,13)]
displayedColNames <- colnames(allDataDf)
colnames(allDataDf) <- c(displayedColNames[1:12],'detail totals')
kable(allDataDf, booktabs = TRUE) %>% kable_styling(font_size = 10)

```

Table 1 shows the basic subject split between Bayesian modeling and deep learning with total for the two categories, following are the individual package counts with a 'detail totals' as well for the individual packages. Note that the totals do not match because an article can match more than one package and as result gets contributes to the count of multiple packages. On average an article's journal is classified into two subject categories.

Staring with the subject that got us here we see that computer science has 728 Bayesian publications vs 29,301 deep learning publications for 98% of the share of 30,029 total. The left most two columns are the same queries used for the graph in Fig 1 with the total in column 3. Columns 4-12 are individual packages with percentage of the total in the 13th column.  Sticking with computer science, one sees that TensorFlow has 18,178 references with 51% of the count--note that the total counts are higher because an article can matched by more that one query, each package gets a query, so the article count is 3 if Keras, PyMC and TensorFlow queries match the references and it has a computer science subject classification.

I leave the tea leaf interpretation to others but I'll make the observation that Bayesians are playing a pretty big game in a lot of areas of science and as someone interested in getting funding to Bayesian software this is a pretty strong argument that we are highly relevant.

# Details of the queries

Below are the queries used to identify the packages. Direct citation linking does not work well with these packages since many of them had no journal publication to cite in a referring article. As a result a string search was used to identify the citations in the reference section as determined by Scopus.com's indexing engine. I would estimate a 5% false positive rate based based on informal examination of return sets.

For an example of how the queries were developed, the stand alone query 'rstan' had the addition of rejecting articles that mentioned 'mit' because there was a common reference to 'http://wwwmath.mit.edu/~rstan/ec/' that existed prior to 2012 when Stan was released. The name 'Stan' presented obvious difficulties and the combination shown achieved desired count drop off at 2012 with reasonable precision and each of the 'OR' terms contributed to coverage significantly. How the word 'mc-stan.org' was tokenized is unknown but there were no results before 2013 for the solo query.

```{r Query Table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
pkg_query_m_display <- pkg_query_m
pkg_query_m_display[,2] <- sprintf("REF(%s)", str_replace_all(pkg_query_m[,2], '\\+',' '))

kable(pkg_query_m_display, booktabs = TRUE) %>% kable_styling(font_size = 12)

```

PyMC presented some challenges in that the string PyMC was referenced in early publications but also is ambiguous with chemical terms, so like Stan, author names were included to restrict search as was the case with 'brms'. These queries were initially developed to search entire documents rather than the current restriction to the references section of papers. I would expect the false positive rate to be even lower.

# Access to source/scripts and acknowledgements

Supporting technology for this document includes the R language (@Rcite), the 'tidyverse' group of packages (@tidyverse) that give data and visualization tools and this Rmarkdown document (@rmarkdown) rendered with 'bookdown' (@bookdown) developed using the Rstudio (@rstudio) IDE (Integrated Development Environment). Supporting package include 'kable-extra' (@kable-extra), a table formatter, the 'scales' (@scales) package for converting to percentages, 'ggrepel' (@ggrepel) for labeling line graphs with dynamically shifting labels, 'jsonlite' (@jsonlite) for JSON data parsing, 'httr' (@httr) for GET requests, 'stringr' (@stringr) for regular expression matching and 'redux' (@redux) for access to the Redis server (@redis) which I highly recomend using as a caching layer for this sort of project.

I'd like to thank the Andrew Gelman's Laplace lab at Columbia University for support.

I intend this to be an evolving document with periodic updates at it's github repo at [https://github.com/breckbaldwin/bayesianImpactMetrics](https://github.com/breckbaldwin/bayesianImpactMetrics) where it lives as `DeepLearningAndBayesianSoftware.Rmd`, is rendered as `DeepLearningAndBayesianSoftware.html` and is viewable as [https://breckbaldwin.github.io/bayesianImpactMetrics/DeepLearningAndBayesianSoftware.html](https://breckbaldwin.github.io/bayesianImpactMetrics/DeepLearningAndBayesianSoftware.html). All source code is viewable in the .Rmd document.

I can be reached at breckbaldwin@gmail.com or via the github issues on the repository.

# References